---
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  hide
params:
  ReportTitle:    "Kansas Primary Election 2020-11-03"
  ElectionName:   "General"
  ElectionDate:   "2020-11-03"
  AfterFileDate:  "2021-02-05"
  HistoryIndex:   5
  District:       "Statewide"
  DistrictName:   "Kansas"
  AgeLimit:       105

title:  "`r params$ReportTitle`"
author: "Earl F Glynn"
date: "<small>`r Sys.Date()`</small>"
---

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r setup2, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment    = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

##  {.active}

## Constants

```{r}
EXCEL_LIMIT <- 2^20
```

```{r}
params
```

## Packages

```{r}
library(tidyverse)
library(lubridate)   # date functions
library(scales)      # comma

library(readxl)
library(writexl)
library(kableExtra)

library(gplots)      # heatmap.2
```

## ggplot theme

```{r}
theme_set(theme_minimal() +

          theme(axis.text             = element_text(size = 10),
                axis.title            = element_text(size = 14),

                legend.position       = "bottom",

                plot.caption          = element_text(hjust = c(0.0,1.0),
                                                     size = 10),
                plot.caption.position = "plot",

                plot.title.position   = "plot",

                strip.background      = element_rect(fill = "aliceblue"),
                strip.text            = element_text(size = 14),

                title                 = element_text(size = 14)))
```

## Helper Functions

```{r}
Show <- function(data, caption="", bigMark="", ...)
{
  data                                                    %>%
  kable("html",
        caption = caption,
        format.args = list(big.mark = bigMark))   %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"),
                position = "left",
                full_width = FALSE, ...)
}
```

Read `efg`s "standardized" preprocessed version of the voter file from the Kansas Secretary of State.

```{r}
readVoterFile <- function(Filename)
{
  voterFile <-
  read_csv(Filename, guess_max = EXCEL_LIMIT) %>%
  select(
          ID14,                   # Hash of County, Name, DOB, Gender
          SOSKey,
          Status,
          BirthDate,
          RegisterDate,
          Gender,
          County, City,
          Precinct, Precinct2,
          Party,                 # D, R, L, U, ?
          LastVoteYear,
          TotalVotes,            # Number of ballots in  voting history (max is 10)
          August,                # Number of August ballots cast in even years
          AugPattern,            # Right most "X" is for most recent August Primary
          November,              # Number of November ballots cast in even years
          NovPattern             # Right most "X" is for most recent November General
        )             %>%
  mutate(
          Gender       = recode(Gender,
                                "F"      = "Female",
                                "M"      = "Male",
                                "?"      = "Unknown",
                                .missing = "Unknown"),
          Party        = recode(Party,
                                "?"      = "U",
                                .missing = "U"),
          BirthDate    = as.Date(BirthDate,    "%m/%d/%Y"),
          RegisterDate = as.Date(RegisterDate, "%m/%d/%Y"),
          Age          = round((BirthDate %--% as.Date(ELECTION_DATE)) /
                               years(1)),
          Registered   = (RegisterDate %--% as.Date(ELECTION_DATE)) /
                          years(1),
          RegisteredTooLate = as.integer(
                                          as.Date(RegisterDate, "%m/%d/%Y") >
                                          (as.Date(ELECTION_DATE) - 21)
                                         )
        )

  invisible(voterFile)
}
```                                 

```{r}
ageDataByGroup <- function(VOTERS, GROUP)
{
  VOTERS                   %>%
  group_by(Age, {{GROUP}}) %>%
  summarize(Registered     = n(),
            Active         = sum(Status == "A",   na.rm = TRUE),
            Voted          = sum(VotedInElection, na.rm = TRUE),

            ActivePercent  = 100 * Active / Registered,
            VotedPercent   = 100 * Voted  / Registered,
            .groups = "drop")   %>%
  ungroup()
}
```

```{r}
bySummary <- function(VOTERS, GROUP)
{
  VOTERS               %>%
  group_by({{GROUP}})  %>%
  summarize(Registered = n(),
            Active     = sum(Status == "A",   na.rm = TRUE),
            Voted      = sum(VotedInElection, na.rm = TRUE),

            ActivePercent = 100 * Active / Registered,
            VotedPercent  = 100 * Voted  / Registered,
            .groups = "drop")   %>%
  ungroup()
}
```

# Parameters

Use first publicly available voter list after a general election whenever possible.

```{r}
ELECTION_DATE    <- params$ElectionDate   # Use for age computations   
AFTER_FILE_DATE  <- params$AfterFileDate                                 
HISTORY_INDEX    <- params$HistoryIndex       
ELECTION_NAME    <- params$ElectionName
```

*Data for `r ELECTION_NAME` Election `r ELECTION_DATE` is from file `r AFTER_FILE_DATE`.*

Registration for Kansas elections closes 21 days before an election. See [Voter Registration Deadlines](https://www.ncsl.org/research/elections-and-campaigns/voter-registration-deadlines.aspx).

Voters registered after that date are excluded from analysis, unless voter history indicates they voted. There’s no practical way to resolve such inconsistent data.

```{r}
FigureDistrict <- paste(params$DistrictName, params$District)
FigureTitle    <- paste("Kansas", ELECTION_NAME, "Election", ELECTION_DATE)
Source         <- paste("Source:  Kansas Secretary of State, Voter File ",
                        AFTER_FILE_DATE)
Stamp          <- paste("efg", format(time.1, "%Y-%m-%d %H%M"))
```

# Read Kansas Data

```{r}
afterFilename <- paste0(paste0("E:/Data/Kansas/RawData/",
                               str_sub(ELECTION_DATE, 1, 3), "0s/"),     #####
                        str_sub(AFTER_FILE_DATE, 1, 4), "/",
                        AFTER_FILE_DATE, "/KSCVRID.CSV")
voterList <- readVoterFile(afterFilename) 
nrow(voterList)          
```

## Add `VotedInElection` variable  

```{r}
if (ELECTION_NAME == "Primary")
{ 
  voterList <-
    voterList %>%
    mutate(
            VotedInElection = 
              as.integer(substr(AugPattern, HISTORY_INDEX, HISTORY_INDEX) == "X")
          )
} 

if (ELECTION_NAME == "General")
{
  voterList <-
    voterList %>%
    mutate(
            VotedInElection = 
              as.integer(substr(NovPattern, HISTORY_INDEX, HISTORY_INDEX) == "X")
          )
}
```

## Exclude if registered after 3 weeks before election

```{r}
table(voterList$VotedInElection, voterList$RegisteredTooLate, useNA = "ifany")
```

```{r}
registeredTooLate <-
  voterList  %>%
  filter(VotedInElection   == 0,
         RegisteredTooLate == 1)

nrow(registeredTooLate)
```

```{r}
voterList <-
  voterList   %>%
  filter(VotedInElection   == 1  |
         RegisteredTooLate == 0)

nrow(voterList)
```

# Statewide Summary

```{r}
voterList %>%
  summarize(
             nRecords = n(),
             nID14    = n_distinct(ID14),
             nSOSKey  = n_distinct(SOSKey),
             nVoted   = sum(VotedInElection, na.rm = TRUE),
             nParty   = n_distinct(Party),
             nGender  = n_distinct(Gender),
             nCounty  = n_distinct(County),
             nPrecinct = n_distinct(Precinct),
             minAge   = min(Age, na.rm = TRUE),
             maxAge   = max(Age, na.rm = TRUE)
            )  %>%
  Show(bigMark = ",")
```

When `nSOSKey` > `nID14` there are very likely some duplicate voter records that should be deleted.

Number of voters with age > 105 years old (or missing birthdates):

```{r}
tooOld <-
  voterList  %>%
  filter(Age > params$AgeLimit | is.na(Age))

nrow(tooOld)
```

These "too old" voters (or those without age) written to file `Too-Old-Voters.xlsx'

```{r}
if (nrow(tooOld) > 0)
{
  write_xlsx(tooOld, paste0("Too-Old-Voters-", params$DistrictName, ".xlsx"))
}
```

Any too young?

```{r}
tooYoung <-
  voterList  %>%
  filter(Age < 18)

nrow(tooYoung)
```

```{r}
if (nrow(tooYoung) > 0)
{
  write_xlsx(tooYoung, paste0("Too-Young-Voters-", params$DistrictName, ".xlsx"))
}
```

Voter count after removal of "too young", "too old" and missing age

```{r}
voterList <-
  voterList  %>%
  filter(Age >= 18,
         Age <= params$AgeLimit,
         !is.na(Age))

nrow(voterList)
```

## Counts by Age

Number of age intervals

```{r}
ageData  <-
  voterList            %>%
  group_by(Age)        %>%
  summarize(Registered = n(),
            Active     = sum(Status == "A",   na.rm = TRUE),
            Voted      = sum(VotedInElection, na.rm = TRUE),

            ActivePercent = 100 * Active / Registered,
            VotedPercent  = 100 * Voted  / Registered,
            .groups = "drop")   %>%
  ungroup()

nrow(ageData)
```

Total number of Registered, Active, Ballots Cast

```{r}
countsRegistered <- sum(ageData$Registered, na.rm = TRUE)
countsActive     <- sum(ageData$Active,     na.rm = TRUE)
countsVoted      <- sum(ageData$Voted,      na.rm = TRUE)

c(countsRegistered, countsActive, countsVoted)
```

## Voter Counts: Registered, Active, Voted by Age

```{r}
countPlotData <-
  ageData                                                       %>%
  select(Age,
         `Registered Voters` = Registered,
         `Active Voters`     = Active,
         `Ballots Cast`      = Voted)                           %>%
  pivot_longer(-Age, names_to = "Group", values_to = "Voters")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Registered Voters",
                                   "Active Voters",
                                   "Ballots Cast")))            %>%
  arrange(Group, Age)
```

```{r RegisteredActiveVoted, fig.width = 8, fig.height = 6}
registeredActiveVoted <-
  ggplot(countPlotData, aes(Age, y= Voters, color = Group)) +
  geom_line(size = 2)                                    +
  scale_y_continuous(labels = scales::comma)             +
  scale_x_continuous(breaks = 20 + 0:8 * 10)             +
  scale_color_manual(values = c("grey", "black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  Voters by Age",
                       " - ", FigureDistrict),
        subtitle = paste0(comma(countsRegistered), " Registered Voters; ",
                          comma(countsActive),     " Active Voters; ",
                          comma(countsVoted),      " Ballots Cast "),
        x = "Election Day Age[years]",
        y = "Number of Voters by Age",
        caption = c(Source, Stamp)
      )

print(registeredActiveVoted)
```

```{r}
write_xlsx(countPlotData %>% spread(Group, Voters),
           "Voter-Counts-by-Age.xlsx")
```


## Fraction of Registered Voters by Age: Active, Voted 

```{r}
fractionPlotData <-
  ageData                                                            %>%
  select(Age,
         `Active Voters` = ActivePercent,
         `Ballots Cast`  = VotedPercent)                              %>%
  pivot_longer(-Age, names_to = "Group", values_to = "Percent")       %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voters", "Ballots Cast"))) %>%
  arrange(Group, Age)
```

```{r}
overallFractionActive = sum(ageData$Active) / sum(ageData$Registered)
overallFractionVoted  = sum(ageData$Voted)  / sum(ageData$Registered)

c(overallFractionActive, overallFractionVoted)
```

```{r ActiveVotedPercentage, fig.width = 8, fig.height = 6}
activeVoted <-
  ggplot(fractionPlotData, aes(Age, y= Percent, color = Group))    +
  geom_line(size = 2)                                    +

  geom_hline(yintercept = 100 *overallFractionActive,
             color = "black",
             linetype = "dotted", size = 1)              +
  geom_text(aes(x = 15,
                y = 100 * overallFractionActive,
                label = paste0(round(100 * overallFractionActive, 2), "%")),
                color = "black", hjust = 0, vjust = 1, size = 4)   +

  geom_hline(yintercept = 100 *overallFractionVoted,
             color = "red",
             linetype = "dotted", size = 1)                        +
  geom_text(aes(x = 15,
                y = 100 * overallFractionVoted,
                label = paste0(round(100 * overallFractionVoted, 2), "%")),
                color = "red", hjust = 0, vjust = 1, size = 4)     +

  scale_y_continuous(breaks =      0:10 * 10)            +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)            +
  scale_color_manual(values = c("black", "red"))         +
  labs(
        title = paste0(FigureTitle, ":  Percent of Registered Voters",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Percentage of Registered Voters",
        caption = c(Source, Stamp)
      )                                                  +
  theme(title = element_text(size = 12))

print(activeVoted)
```
```{r}
write_xlsx(fractionPlotData %>% spread(Group, Percent),
           "Voter-Percentages-by-Age.xlsx")
```

## Normalized Fraction

The normalization is formed by the ratio of turnout by age to the overall turnout for all ages (the dotted line in the diagram above).

What is the intuition for this normalization?

* Normalized fraction < 1 shows ages that "underperformed".

* Normalized fraction > 1 shows ages that "outperformed".

* Normalized fraction can be compared across different districts, like state and county.

```{r}
normalized <-
  ageData    %>%
  mutate(ActiveNormalized = (ActivePercent/100) / overallFractionActive,
         VotedNormalized  = (VotedPercent/100)  / overallFractionVoted)
```

```{r}
normalizedPlotData <-
  normalized                                                          %>%
  select(Age,
         `Active Voters` = ActiveNormalized,
         `Ballots Cast`  = VotedNormalized)                           %>%
  pivot_longer(-Age, names_to = "Group", values_to = "Ratio")         %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voters", "Ballots Cast"))) %>%
  arrange(Group, Age)
```

```{r NormalizedActiveVoted, fig.width = 8, fig.height = 6}
normalizedVoted <-
  ggplot(normalizedPlotData, aes(Age, y= Ratio, color = Group))  +
  geom_line(size = 2)                                            +
  geom_hline(yintercept = 1.0, color = "blue",
             linetype = "dotted", size = 1)                      +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)                    +
  scale_y_continuous(breaks = 0.01 * 0:24 * 10)                  +
  scale_color_manual(values = c("black", "red"))                 +
  labs(
        title = paste0(FigureTitle, ":  Normalized Voter Fraction",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Normalized Fraction of Voters",
        caption = c(Source, Stamp)
      )                                                          +
  theme(title = element_text(size = 12))

print(normalizedVoted)
```
```{r}
write_xlsx(normalizedPlotData %>% spread(Group, Ratio),
           "Voter-Normalized-by-Age.xlsx")
```

# By Party

```{r}
ageDataByParty <- ageDataByGroup(voterList, Party)
nrow(ageDataByParty)
```

```{r}
byParty <- bySummary(voterList, Party)
byParty %>%  Show(bigMark = ",")
```

```{r}
countPlotDataByParty <-
  ageDataByParty                                                          %>%
  select(Age, Party,
         `Registered Voters` = Registered,
         `Active Voters`     = Active,
         `Ballots Cast`      = Voted)                                     %>% 
  pivot_longer(c(-Age,-Party), names_to = "Group", values_to = "Voters")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Registered Voters",
                                   "Active Voters",
                                   "Ballots Cast")))                      %>%
  arrange(Group, Age, Party)

nrow(countPlotDataByParty)
```

## Voter Registration Counts

```{r}
partyCounts <- table(voterList$Party)
partyCounts
```


```{r}
partyLabel <- c("D" = paste("Democratic",   comma(partyCounts["D"])),
                "L" = paste("Libertarian",  comma(partyCounts["L"])),
                "R" = paste("Republican",   comma(partyCounts["R"])),
                "U" = paste("Unaffiliated", comma(partyCounts["U"])))
```

```{r}
if (length(partyCounts) == 5)   # 2010 and earlier
{
  partyLabel = c(partyLabel[1],
                 "F" = paste("Reform",  comma(partyCounts["F"])),
                 partyLabel[2:4])
}
```

```{r}
registeredActiveVotedByParty <-
  ggplot(countPlotDataByParty,
         aes(x = Age, y= Voters, color = Group))         +
  geom_line(size = 2)                                    +
  scale_y_continuous(labels = scales::comma)             +
  scale_x_continuous(breaks = 20 + 0:8 * 10)             +
  scale_color_manual(values = c("grey", "black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  Voters by Party and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Number of Voters by Age",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank())            +
  facet_wrap(~Party,
             ncol = length(partyLabel),
             labeller = labeller(Party = partyLabel))    +
  geom_text(data = byParty,
            aes(x = 15,
                y = Inf,
                label = paste0(comma(Active), " Active Voters\n",
                               comma(Voted),  " Ballots Cast")),
            color = "black", hjust = 0, vjust = 1, size = 4)
```

```{r RegisteredActiveVotedByParty, fig.width = 12, fig.height = 6}
print(registeredActiveVotedByParty)
```

## Fraction of Voters

```{r}
partyLabelFraction <-
  c("D" = paste0("Democratic ",
                 round(100 * partyCounts["D"] / sum(partyCounts), 2), "%"),

    "L" = paste0("Libertarian ",
                 round(100 * partyCounts["L"] / sum(partyCounts), 2), "%"),

    "R" = paste0("Republican ",
                 round(100 * partyCounts["R"] / sum(partyCounts), 2), "%"),

    "U" = paste0("Unaffiliated ",
                 round(100 * partyCounts["U"] / sum(partyCounts), 2), "%"))
```

```{r}
if (length(partyCounts) == 5)   # 2010 and earlier
{
  partyLabelFraction = c(partyLabelFraction[1],
                         "F" = paste0("Reform ",
                                      round(100 * partyCounts["F"] / sum(partyCounts), 2), "%"),
                         partyLabelFraction[2:4])
}
```

```{r}
fractionByPartyAge <-
  ageDataByParty                                                      %>%
  select(Age, Party,
         `Active Voters` = ActivePercent,
         `Ballots Cast`  = VotedPercent)                              %>%
  pivot_longer(c(-Age, -Party), names_to = "Group", values_to = "Percent")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voters", "Ballots Cast"))) %>%
  arrange(Group, Age, Party)
```

```{r}
fractionByParty <-
  ageDataByParty
```

```{r}
activeVotedByParty <-
  ggplot(fractionByPartyAge, aes(Age, y= Percent, color = Group)) +
  geom_line(size = 2)                                    +
  scale_y_continuous(breaks =      0:10 * 10)            +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)            +
  scale_color_manual(values = c("black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  Percentage of Registered Voters by Party and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Percentage of Registered Voters",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())            +

  facet_grid(col = vars(Party),
             labeller = labeller(Party = partyLabelFraction),
             scales = "free_x")                          +

  geom_hline(data = byParty,
             aes(yintercept = ActivePercent),
             color = "black",
             linetype = "dotted", size = 1)              +
  geom_text(data = byParty,
            aes(x = 15,
                y = ActivePercent,
                label = paste0(round(ActivePercent, 1), "%")),
            color = "black", hjust = 0, vjust = 1, size = 4)   +

  geom_hline(data = byParty,
             aes(yintercept = VotedPercent),
             color = "red",
             linetype = "dotted", size = 1)               +
  geom_text(data = byParty,
            aes(x = 15,
                y = VotedPercent,
                label = paste0(round(VotedPercent, 1), "%")),
            color = "red", hjust = 0, vjust = 1, size = 4)
```

```{r ActiveVotedPercentageByParty, fig.width = 12, fig.height = 6}
print(activeVotedByParty)
```

# By Gender

```{r}
ageDataByGender <- ageDataByGroup(voterList, Gender)
nrow(ageDataByGender)
```

```{r}
byGender <- bySummary(voterList, Gender)
byGender %>% Show()
```

```{r}
countPlotDataByGender <-
  ageDataByGender                             %>%
  select(Age, Gender,
         `Registered Voters` = Registered,
         `Active Voters`     = Active,
         `Ballots Cast`      = Voted)         %>%
  pivot_longer(c(-Age,-Gender), names_to = "Group", values_to = "Voters")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Registered Voters",
                                   "Active Voters",
                                   "Ballots Cast"))) %>%
  arrange(Group, Age, Gender)

nrow(countPlotDataByGender)
```

## Voter Registration Counts

```{r}
genderCounts <- table(voterList$Gender)
genderCounts
```


```{r}
genderLabel <- c("Female"  = paste("Female",  comma(genderCounts["Female"])),
                 "Male"    = paste("Male",    comma(genderCounts["Male"])),
                 "Unknown" = paste("Unknown", comma(genderCounts["Unknown"])))
```

```{r}
registeredActiveVotedByGender <-
  ggplot(countPlotDataByGender,
         aes(x = Age, y= Voters, color = Group))         +
  geom_line(size = 2)                                    +
  scale_y_continuous(labels = scales::comma)             +
  scale_x_continuous(breaks = 20 + 0:8 * 10)             +
  scale_color_manual(values = c("grey", "black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  By Gender and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Number of Voters by Age",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank())            +
  facet_wrap(~Gender,
             ncol = length(genderLabel),
             labeller = labeller(Gender = genderLabel))    +
  geom_text(data = byGender,
            aes(x = 15,
                y = Inf,
                label = paste0(comma(Active), " active voters\n",
                               comma(Voted),  " ballots cast")),
            color = "black", hjust = 0, vjust = 1, size = 4)
```

```{r RegisteredActiveVotedByGender, fig.width = 12, fig.height = 6}
print(registeredActiveVotedByGender)
```

## Fraction of Voters

```{r}
genderLabelFraction <-
  c("Female" = paste0("Female ",
                 round(100 * genderCounts["Female"] / sum(genderCounts), 2), "%"),

    "Male" = paste0("Male ",
                 round(100 * genderCounts["Male"]    / sum(genderCounts), 2), "%"),

    "Unknown" = paste0("Unknown ",
                 round(100 * genderCounts["Unknown"] / sum(genderCounts), 2), "%"))
```

```{r}
fractionByGenderAge <-
  ageDataByGender                                                     %>%
  select(Age, Gender,
         `Active Voters` = ActivePercent,
         `Ballots Cast`  = VotedPercent)                              %>%
  pivot_longer(c(-Age, -Gender), names_to = "Group", values_to = "Percent")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voters", "Ballots Cast"))) %>%
  arrange(Group, Age, Gender)
```

```{r}
fractionByGender <-
  ageDataByGender
```

```{r}
activeVotedByGender <-
  ggplot(fractionByGenderAge, aes(Age, y= Percent, color = Group)) +
  geom_line(size = 2)                                    +
  scale_y_continuous(breaks =      0:10 * 10)            +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)            +
  scale_color_manual(values = c("black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  Percentage By Gender and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Percentage of Registered Voters",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())            +

  facet_grid(col = vars(Gender),
             labeller = labeller(Gender = genderLabelFraction),
             scales = "free_x")                          +

  geom_hline(data = byGender,
             aes(yintercept = ActivePercent),
             color = "black",
             linetype = "dotted", size = 1)              +
  geom_text(data = byGender,
            aes(x = 15,
                y = ActivePercent,
                label = paste0(round(ActivePercent, 1), "%")),
            color = "black", hjust = 0, vjust = 1, size = 4)   +

  geom_hline(data = byGender,
             aes(yintercept = VotedPercent),
             color = "red",
             linetype = "dotted", size = 1)               +
  geom_text(data = byGender,
            aes(x = 15,
                y = VotedPercent,
                label = paste0(round(VotedPercent, 1), "%")),
            color = "red", hjust = 0, vjust = 1, size = 4)
```

```{r ActiveVotedPercentageByGender, fig.width = 12, fig.height = 6}
print(activeVotedByGender)
```

# By County

```{r}
ageDataByCounty <- ageDataByGroup(voterList, County)
nrow(ageDataByCounty)
```

```{r}
byCounty <-
  bySummary(voterList, County)  %>%
  rename(OverallActivePercent = ActivePercent,
         OverallVotedPercent  = VotedPercent)

byCounty                %>%  
  arrange(-Registered)  %>%
  Show(bigMark = ",")   %>%  
  scroll_box(height = "300px")
```

## County Size

```{r County_Density}
options(scipen = 999)
ggplot(byCounty, aes(x = Registered))                           +
  geom_histogram(fill = "skyblue", color = "grey80", bins = 50) +
  scale_x_continuous(labels = scales::comma)                    +
  labs(title = paste("County Sizes in", FigureDistrict),
       subtitle = paste("All", nrow(byCounty), "counties"),
       x = "County Size [registered voters]",
       y = "Count",
       caption=c(Source, Stamp))
```

For now, let's only look at voter turnout.

```{r}
normalizedAgeDataByCounty <-
  byCounty                                     %>%
  select(County, OverallVotedPercent)          %>%
  inner_join(ageDataByCounty,
             by = "County")                    %>%
  mutate(NormalizedFractionVoted = VotedPercent / OverallVotedPercent)  %>%
  select(County, Age, NormalizedFractionVoted) %>%
  arrange(County, Age)

dim(normalizedAgeDataByCounty)
```

## Voter Registration Counts

```{r}
countyCounts <- table(voterList$County)
```

```{r}
sum(countyCounts)
```

## Scatterplot of State Overall with All Counties

```{r}
stateNormalized <-
  normalizedPlotData                       %>%
  rename(NormalizedFractionVoted = Ratio)  %>%
  filter(Group == "Ballots Cast")          %>%
  select(-Group)                           %>%
  mutate(County = "Kansas")      %>%
  relocate(County, .before = Age)

dim(stateNormalized)
```

```{r NormalizedVotedTurnout, fig.width = 8, fig.height = 6}
normalizedVoted <-
  ggplot()                                      +

  geom_line(data = normalizedAgeDataByCounty,
            aes(x = Age, y= NormalizedFractionVoted,
                group = County),
            color = "grey", size = 1)           +

  geom_line(data = stateNormalized,
            aes(x = Age, y= NormalizedFractionVoted),
            color = "red", size = 2)            +

  geom_hline(yintercept = 1.0, color = "blue",
             linetype = "dotted", size = 1)     +

  geom_text(aes(x = 85,
                y = 0.1,
                label = "Low counts cause 'noise' for older ages"),
            color = "black", hjust = 1, vjust = 1, size = 4)   +

  scale_x_continuous(breaks = 2:10 * 10)        +
  scale_y_continuous(breaks = 0:12 *  0.2,      
                     limits = c(0, 2.5) )       +

  labs(
        title = paste0(FigureTitle, ":  Normalized Voter Faction by Age"),
        subtitle = paste0(FigureDistrict, " = red line; ",
                          length(countyCounts), " Counties = grey lines;"),
        x = "Election Day Age[years]",
        y = "Normalized Fraction of Voters",
        caption = c(Source, Stamp)
      )                                      +
  theme(
         panel.grid.minor.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         title = element_text(size = 12)
       )

print(normalizedVoted)
```

## Heatmap

Create matrix

```{r}
Kansas <-
  bind_rows(stateNormalized,
            normalizedAgeDataByCounty)  %>%
  pivot_wider(names_from = Age,
              values_from = NormalizedFractionVoted,
              values_fill = list(NormalizedFractionVoted = 0))  %>%
  mutate(across(everything(), ~replace_na(.x, 0)))   # kludge fix (e.g., Primary 2006)

```

```{r}
comparisonMatrix <- as.matrix(Kansas %>% select(-County))
dim(comparisonMatrix)
```

```{r}
dimnames(comparisonMatrix) <- list(
                                    Kansas$County,
                                    colnames(Kansas[-1]) %>% as.numeric()
                                   )
```

```{r}
comparisonMatrix[1:5, 1:5]
```

```{r}
ratioStats <- fivenum(comparisonMatrix)

ratioStats
```

```{r}
colorBreaks <-
  c(
    seq(0.0, 1.0, length.out = 9),
    seq(1.0, 2.0, length.out = 9)[-1]
  )

colorBreaks
```

Attempt to "balance" palette above and below 1.

```{r}
colorPalette <-
  c(colorRampPalette(c("red",   "white"))(9)[1:8],    # skip last
    "#FFFFFF",                                        # color for ratio of 1
    colorRampPalette(c("white", "skyblue"))(8)[2:8])  # skip first

colorPalette
```

```{r}
# layout matrix 1 = heatmap, 2 = row dendrogrm, 3 = column dendrogram, 4 = key
lmat <- rbind( c(0, 3, 3),  # layout matrix
               c(2, 1, 1),
               c(0, 0, 4))

lwid <- c(3, 4, 4)          # entry for each column in layout matrix

# relative height of the rows. must have length = nrow(lmat)
lhei <- c(0.25, 4, 0.20)    # entry for each row in layout matrix
```

```{r heatmap, fig.width = 20, fig.height = 32}
oldPar <- par(mar = c(5, 0, 2, 0))

heatmap.2(
           comparisonMatrix,
           scale      = "none",
           dendrogram = "row",
           Colv       = NA,         # suppress reordering of columns

           trace  = "none",         # turns off trace lines inside the heat map
           col    = colorPalette,
           breaks = colorBreaks,

           margins = c(4, 10),      # column and row names

           cexRow = 1.5,
           cexCol = 1.25,

           key.title = NA,           # suppress title
           key.xlab  = paste0("Normalized Voter Fraction ", ELECTION_DATE),
           keysize   = 3,
           density.info = "none",    # turns off density plot inside color legend

           lmat = lmat, lwid = lwid, lhei = lhei
         )
title(main = paste0(FigureTitle, 
                    ":  Normalized Voter Faction by Age"), 
      cex.main = 3,
      sub = paste(Source, " - ", Stamp),
      cex.sub  = 1.25)

par(oldPar)
```

# Fini {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r}
sessionInfo()
```

</div>

```{r ThatsAll, echo = FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

