---
title:  "Driver to Create Statewide Normalized Turnout Reports"
author: "Earl F Glynn"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

```{r Dsetup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r Dsetup2, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment    = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

##  {.active}

## Constants

```{r DConstants}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r DPackages}
library(rmarkdown)
library(tidyverse)
library(lubridate)

library(readxl)
library(kableExtra) 
```

## Helper Functions

```{r DHelper}
Show <- function(data, caption="", bigMark="", ...)
{
  data                                                    %>%
  kable("html", 
        caption = caption,
        format.args = list(big.mark = bigMark))   %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"),
                position = "left", 
                full_width = FALSE, ...)
}
```

# Metadata

```{r DShow}
metadata <- read_xlsx("Statewide-Elections.xlsx") 

metadata  %>% Show()
```

# Function to Process Election

```{r Dfunction}
processElection <- function(info)
{
  title <- paste0("Kansas ", info$ELECTION_NAME, " Election ", info$ELECTION_DATE)    
  
  OUTPUT_DIR <- paste0("./", info$ELECTION_DATE)      

  rmarkdown::render(
                     "Statewide-Normalized-Turnout-Template.Rmd",
                     output_format = html_document(theme      = "cerulean",
                                                   toc        = TRUE,
                                                   toc_format = TRUE,
                                                   toc_float  = TRUE,
                                                   number_sections = TRUE,
                                                   code_folding = "hide"),

                     output_file = paste0(info$ELECTION_DATE, 
                                          "-Statewide-Normalized-Turnout.html"),
                     output_dir        = OUTPUT_DIR,

                     quiet  = TRUE,
                     params = list(
                                    ReportTitle     = title,
                                    ElectionDate    = info$ELECTION_DATE,
                                    ElectionName    = info$ELECTION_NAME,
                                    AfterFileDate   = info$AFTER_FILE_DATE,
                                    HistoryIndex    = info$HISTORY_INDEX
                                  )
                   )

  # Add election name to all graphics files
  listFiles <- dir(path = ".", 
                   pattern = "^.*\\.png$|Voter.*\\.xlsx$")
  newNames <- paste0(info$ELECTION_DATE, "-", listFiles)
  file.rename(listFiles, newNames)

  #Brute force to get all files into `output_dir`
  listFiles <- dir(path = ".", 
                   pattern = "^.*\\.png$|Voter.*\\.xlsx$")
  file.copy(listFiles, OUTPUT_DIR, copy.date = TRUE)
  file.remove(listFiles)   
  
  invisible()
}
```

# Process All Elections

For now, don't bother with parallelizing this.  Markdown writes several files to the same working directory and that's a bit of pain to fix.

```{r Dprocess}
system.time(
             for (electionIndex in 1:nrow(metadata))
             {
               info <- metadata %>% slice(electionIndex)
               cat(info$ELECTION_NAME, as.character(info$ELECTION_DATE), "\n")
               processElection(info)
             }
           )
```

# Fini {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r DSessionInfo}
sessionInfo()
```

</div>

```{r DThatsAllDriver, echo = FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

