---
title: "Ohio Key Heatmap"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

Analysis of data in Ohio file `key.json` from results in Jupyter notebook `Election-Integrity/Ohio/rlee32/Analysis/Ohio-Analysis-4-Generate-Key.html`

combined with county key data from Jupyter notebook
`Ohio-Analysis-3-All-Counties.html`

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r Constants}
FILE_PREFIX   <- "x"    # no spaces
```

```{r}
EXCEL_LIMIT <- 2^20
```

```{r}
stamp <- paste("efg", format(time.1, "%Y-%m-%d %H%M"))
```

## Packages

```{r Packages}
library(tidyverse)
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

```{r}
library(RJSONIO)
```

 Graphics

```{r Graphics}
library(scales)      # comma, comma_format
```

## ggplot theme

[ggplot Themes Gallery](https://www.datanovia.com/en/blog/ggplot-themes-gallery/)

```{r ggplotTheme}
theme_set(theme_minimal() +

          theme(axis.text             = element_text(size = 10),
                axis.title            = element_text(size = 14),

                plot.caption          = element_text(hjust = c(0.0,1.0)),
                plot.caption.position = "plot",   # R 4.0.0

                plot.title.position   = "plot",
                title                 = element_text(size = 14))
           )

COLOR_BAR     <- "skyblue"
COLOR_OUTLINE <- "grey80"
```

## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       %>%
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  %>%
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       %>%
  scroll_box(height = height, width = width) 
}
```

# Read `Key` data

## All Counties

```{r}
OhioNormalized <- 
  read_csv("Ohio-County-Normalized-Turnout-by-Age.csv")  %>%
  select(-c(Age101:Age105))

dim(OhioNormalized)
```

## Ohio statewide

```{r}
JSON <- fromJSON("key.json")

OhioKeyRaw <- 
  tibble(age = as.numeric( names(JSON) ),
         keyValue = as.numeric(JSON))  %>%
  arrange(age)

dim(OhioKeyRaw)
```

Make sure there are no gaps in age

Restrict to age 100 to match Dr. Frank's plot as closely as possible.

```{r}
OhioKeyRaw <-
  OhioKeyRaw  %>%
  filter(age <= 100)

nrow(OhioKeyRaw)
```

## Combine

```{r}
Ohio <-
  bind_rows(OhioNormalized,
            bind_cols(Code = 89, County = "OHIO"))

Ohio[89, 3:85] <-as.list(OhioKeyRaw$keyValue)

Ohio %>%
  tail(2)  %>%
  Show()
```

## "Longer" versions

```{r}
countyLonger <-
  Ohio                                 %>%
  filter(County != "OHIO")             %>%
  select(-Code)                        %>%
  pivot_longer(Age018:Age100,
               names_to = "AgeString",
               values_to = "turnout")  %>%
  mutate(Age = str_sub(AgeString, 4, 6) %>% as.numeric())  %>%
  select(-AgeString)                                       %>%
  filter(turnout > 0)                                      %>%  # drop missing
  arrange(County, Age)

dim(countyLonger)
```

```{r}
countyLonger  %>%
  head(3)    %>%
  Show()
```

```{r}
write_xlsx(countyLonger, "County-Longer.xlsx")
```


```{r}
stateLonger <-
  Ohio                                 %>%
  filter(County == "OHIO")             %>%
  select(-Code)                        %>%
  pivot_longer(Age018:Age100,
               names_to = "AgeString",
               values_to = "turnout")  %>%
  mutate(Age = str_sub(AgeString, 4, 6) %>% as.numeric())  %>%
  select(-AgeString)                                       %>%
  arrange(Age)

dim(stateLonger)
```

```{r}
stateLonger  %>%
  head(3)    %>%
  Show()
```

# Ohio Normalized Voter Fraction by Age

## Composite plot with all counties

```{r OhioNormalized, fig.width = 8, fig.height = 6}
TITLE <- "Ohio Normalized Voter Fraction by Age"

normalizedVoted <-
  ggplot()                                      +
  
  geom_line(data = countyLonger, 
            aes(x = Age, y= turnout, 
                group = County),
            color = "grey", size = 1)           +
  
  geom_line(data = stateLonger, 
            aes(x = Age, y= turnout),
            color = "red", size = 2)            +
  
  scale_x_continuous(breaks = 2:10 * 10,
                     limits = c(18, 100) )      +
  scale_y_continuous(breaks = 0:7 * 0.2,
                     limits = c(0, 1.40) )      +
  labs(
        title = TITLE,
        subtitle = paste0("Ohio state = red line; ",
                          "88 counties = grey lines; ",
                          "no exclusions for low counts cause 'noise' for older ages"),
        x = "Election Day Age[years]",
        y = "Normalized Fraction of Voters",
        caption = c("Source:  Ohio Secretary of State, Voter File, 2022-03-25",
                     stamp)
      )                                          +
  theme(title = element_text(size = 12),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())                        

print(normalizedVoted)
```                   

## Heatmap with all counties

```{r}
library(gplots)
```

```{r}
comparisonMatrix <- as.matrix(Ohio %>% select(-Code, -County))
dim(comparisonMatrix)
```

```{r}
dimnames(comparisonMatrix) <- list(
                                    Ohio$County,
                                    str_sub(colnames(Ohio[c(-1, -2)]), 4, 6) %>% as.numeric()
                                   )
```


```{r}
comparisonMatrix[1:5, 1:5]
```

```{r}
ratioStats <- fivenum(comparisonMatrix)

ratioStats
```

```
colorBreaks <- (0:15) / 15 * (max(ratioStats) - min(ratioStats)) 
colorBreaks
```

Want color palette balanced around 1, but data are skewed a bit on the lower side.  So, let's create
pallete with 10 colors below 1, and 4 above 1 to put middle color exactly at 1.

This gives an imbalanced palette:

```
colorPalette <- colorRampPalette(c("red", "white", "skyblue"))
```

```{r}
colorPalette <-
  c(colorRampPalette(c("red",   "white"))(11)[1:10],  # skip last
    "#FFFFFF",                                        # color for ratio of 1
    colorRampPalette(c("white", "skyblue"))(5)[2:5])  # skip first
    
colorPalette
```


```{r}
# layout matrix 1 = heatmap, 2 = row dendrogrm, 3 = column dendrogram, 4 = key
lmat <- rbind( c(0, 3, 3),  # layout matrix
               c(2, 1, 1),
               c(4, 0, 0))

lwid <- c(3, 4, 4)          # entry for each column in layout matrix

# relative height of the rows. must have length = nrow(lmat)
lhei <- c(0.25, 4, 0.25)          # entry for each row in layout matrix
```

```{r heatmap, fig.width = 20, fig.height = 32}

heatmap.2(
           comparisonMatrix,
           scale      = "none",
           dendrogram = "row",
           Colv       = NA,         # suppress reordering of columns
           
           trace  = "none",         # turns off trace lines inside the heat map
           col    = colorPalette,
      #    breaks = colorBreaks,
           
           margins = c(4, 10),      # column and row names
           
           main   = TITLE,  
           
           cexRow = 1.5,    
           cexCol = 1.25,
           
           key.title = NA,           # suppress title
           key.xlab  = "Normalized Turnout Nov 2020",
           keysize   = 2,
           density.info = "none",    # turns off density plot inside color legend
          
           lmat = lmat, lwid = lwid, lhei = lhei
         )                           
```


# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

