---
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  hide
params:
  ReportTitle:      "Normalized Turnout for Smith County, KS" 
  District:         "County"
  DistrictName:     "Smith" 
  ElectionDate:     "2020-11-03"
  ElectionName:     "General"
  AfterFileDate:    "2021-02-05"
  AgeLimit:         105
  PrecinctMinimum:   25
  voterMaster:        0
    
title:  "`r params$ReportTitle`"        
author: "Earl F Glynn"
date: "<small>`r Sys.Date()`</small>"
---

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r setup2, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment    = NA)

time.1 <- Sys.time()
```
         
# Setup {.tabset .tabset-fade .tabset-pills}

##  {.active}

## Constants

```{r}
EXCEL_LIMIT <- 2^20
```

```{r}
params
```     

## Packages                             

```{r}
library(tidyverse)
library(lubridate)   # date functions
library(scales)      # comma

library(readxl)
library(writexl)
library(kableExtra) 

library(gplots)      # heatmap.2
```

## ggplot theme

```{r}
theme_set(theme_minimal() +

          theme(axis.text             = element_text(size = 10),
                axis.title            = element_text(size = 14),
                
                legend.position       = "bottom",
                
                plot.caption          = element_text(hjust = c(0.0,1.0),
                                                     size = 10),
                plot.caption.position = "plot",

                plot.title.position   = "plot",
                
                strip.background      = element_rect(fill = "aliceblue"),
                strip.text            = element_text(size = 14),        
                
                title                 = element_text(size = 14)))
```

## Helper Functions

```{r}
Show <- function(data, caption="", bigMark="", ...)
{
  data                                                    %>%
  kable("html", 
        caption = caption,
        format.args = list(big.mark = bigMark))   %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"),
                position = "left", 
                full_width = FALSE, ...)
}
```

```{r}
ageDataByGroup <- function(VOTERS, GROUP)
{
  VOTERS                   %>%
  group_by(Age, {{GROUP}}) %>%
  summarize(Registered     = n(),
            Active         = sum(Status == "A",   na.rm = TRUE),
            Voted          = sum(VotedInElection, na.rm = TRUE),

            ActivePercent  = 100 * Active / Registered,
            VotedPercent   = 100 * Voted  / Registered,
            .groups = "drop")   %>%
  ungroup()
}
```                               

```{r}
bySummary <- function(VOTERS, GROUP)
{
  VOTERS               %>%
  group_by({{GROUP}})  %>%
  summarize(Registered = n(),
            Active     = sum(Status == "A",   na.rm = TRUE),
            Voted      = sum(VotedInElection, na.rm = TRUE),

            ActivePercent = 100 * Active / Registered,
            VotedPercent  = 100 * Voted  / Registered,
            .groups = "drop")   %>%
  ungroup()  
}
```

# Parameters

Use first publicly available voter list after a general election whenever possible.

```{r}
ELECTION_DATE    <- params$ElectionDate   # Use for age computations    
AFTER_FILE_DATE  <- params$AfterFileDate 

ELECTION_NAME    <- params$ElectionName                  

AGE_LIMIT        <- params$AgeLimit
PRECINCT_MINIMUM <- params$PrecinctMinimum

voterMaster      <- params$voterMaster
```

```{r}
FigureDistrict <- paste(params$DistrictName, params$District)
FigureTitle    <-  paste("Kansas", ELECTION_NAME, "Election", ELECTION_DATE)
Source         <- paste("Source:  Kansas Secretary of State, Voter File ",
                        AFTER_FILE_DATE)
Stamp          <- paste("efg", format(time.1, "%Y-%m-%d %H%M"))
```

# Define Kansas Data

```{r, results = "asis"}
cat("## Filter to Target", FigureDistrict, "\n")
```

```{r}
if (params$District == "County")
{
  voterList <- 
    voterMaster   %>%
    filter(County == params$DistrictName)
}
```

```{r, results = "asis"}
cat("#", FigureDistrict, "Summary\n")
```

```{r}
voterList %>%
  summarize(
             n         = n(),
             nID14     = n_distinct(ID14),
             nSOSKey   = n_distinct(SOSKey),
             nVoted    = sum(VotedInElection, na.rm = TRUE),
             nParty    = n_distinct(Party),
             nGender   = n_distinct(Gender),
             nPrecinct = n_distinct(Precinct),
             minAge    = min(Age, na.rm = TRUE),
             maxAge    = max(Age, na.rm = TRUE)
            )  %>%
  Show(bigMark = ",")
```

When `nSOSKey` > `nID14` there are very likely some duplicate voter records that should be deleted. 

Number of voters with age > 105 years old:

```{r}
tooOld <-
  voterList  %>%
  filter(Age > AGE_LIMIT | is.na(Age))

nrow(tooOld)
```

These "too old" voters (or those without age) written to file `Too-Old-Voters.xlsx'

```{r}
if (nrow(tooOld) > 0)
{
  write_xlsx(tooOld, paste0("Too-Old-Voters-", params$DistrictName, ".xlsx"))
}
```

Any too young?     

```{r}
tooYoung <-
  voterList  %>%
  filter(Age < 18)

nrow(tooYoung)
```

```{r}
if (nrow(tooYoung) > 0)
{
  write_xlsx(tooYoung, paste0("Too-Young-Voters-", params$DistrictName, ".xlsx"))
}
```                

Voter count after removal of "too young", "too old" and missing age    

```{r}
voterList <-
  voterList  %>%
  filter(Age >= 18,
         Age <= params$AgeLimit,
         !is.na(Age))

nrow(voterList)
```            

## Counts by Age

Number of age intervals

```{r}
ageData  <- 
  voterList            %>%
  group_by(Age)        %>%
  summarize(Registered = n(),
            Active     = sum(Status == "A",   na.rm = TRUE),
            Voted      = sum(VotedInElection, na.rm = TRUE),
            
            ActivePercent = 100 * Active / Registered,
            VotedPercent  = 100 * Voted  / Registered,
            .groups = "drop")   %>%
  ungroup()       

nrow(ageData)
```

Total number of Registered, Active, Voted

```{r}
countsRegistered <- sum(ageData$Registered, na.rm = TRUE)
countsActive     <- sum(ageData$Active,     na.rm = TRUE)
countsVoted      <- sum(ageData$Voted,      na.rm = TRUE)

c(countsRegistered, countsActive, countsVoted)
```

## Voter Counts: Registered, Active, Voted by Age

```{r}
countPlotData <- 
  ageData                                                       %>%
  select(Age, 
         `Registered Voters` = Registered, 
         `Active Voters`     = Active, 
         `Ballots Cast`      = Voted)                           %>%     #####
  pivot_longer(-Age, names_to = "Group", values_to = "Voters")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Registered Voters",
                                   "Active Voters",
                                   "Ballots Cast"))) %>%
  arrange(Group, Age)
```

```{r RegisteredActiveVoted, fig.width = 8, fig.height = 6}       
registeredActiveVoted <-
  ggplot(countPlotData, aes(Age, y= Voters, color = Group)) +
  geom_line(size = 2)                                    +
  scale_y_continuous(labels = scales::comma)             +     
  scale_x_continuous(breaks = 20 + 0:8 * 10)             +
  scale_color_manual(values = c("grey", "black", "red")) + 
  labs(
        title = paste0(FigureTitle, ":  Voters by Age",
                       " - ", FigureDistrict),
        subtitle = paste0(comma(countsRegistered), " Registered Voters; ", 
                          comma(countsActive),     " Active Voters; ",
                          comma(countsVoted),      " Ballots Cast "), 
        x = "Election Day Age[years]",
        y = "Number of Voters by Age",
        caption = c(Source, Stamp)
      )                                                  

print(registeredActiveVoted)
```

## Fraction of Registered Voters by Age: Active, Ballots Cast

```{r}
fractionPlotData <- 
  ageData                                                        %>%
  select(Age, 
         `Active Voters` = ActivePercent, 
         `Ballots Cast`  = VotedPercent)                      %>%
  pivot_longer(-Age, names_to = "Group", values_to = "Percent")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voters", "Ballots Cast"))) %>%
  arrange(Group, Age)
```

```{r}
overallFractionActive = sum(ageData$Active) / sum(ageData$Registered)
overallFractionVoted  = sum(ageData$Voted)  / sum(ageData$Registered)

c(overallFractionActive, overallFractionVoted)
```

```{r ActiveVotedPercentage, fig.width = 8, fig.height = 6}
activeVoted <-
  ggplot(fractionPlotData, aes(Age, y= Percent, color = Group))    +
  geom_line(size = 2)                                    +
  
  geom_hline(yintercept = 100 *overallFractionActive, 
             color = "black", 
             linetype = "dotted", size = 1)              +
  geom_text(aes(x = 15,
                y = 100 * overallFractionActive,
                label = paste0(round(100 * overallFractionActive, 2), "%")),
                color = "black", hjust = 0, vjust = 1, size = 4)   +
  
  geom_hline(yintercept = 100 *overallFractionVoted, 
             color = "red", 
             linetype = "dotted", size = 1)                        +
  geom_text(aes(x = 15,
                y = 100 * overallFractionVoted, 
                label = paste0(round(100 * overallFractionVoted, 2), "%")),
                color = "red", hjust = 0, vjust = 1, size = 4)     +
  
  scale_y_continuous(breaks =      0:10 * 10)            +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)            +
  scale_color_manual(values = c("black", "red"))         +
  labs(
        title = paste0(FigureTitle, ":  Percent of Registered Voters",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Percentage of Registered Voters",
        caption = c(Source, Stamp)
      )                                                  +                          
  theme(title = element_text(size = 12))

print(activeVoted)
```                    

## Normalized Fraction 

The normalization is formed by the ratio of turnout by age to the overall turnout for all ages (the dotted line in the diagram above). 

What is the intuition for this normalization?

* Normalized fraction < 1 shows ages that "underperformed".

* Normalized fraction > 1 shows ages that "outperformed".

* Normalized fraction can be compared across different districts, like state and county.

```{r}
normalized <- 
  ageData    %>%
  mutate(ActiveNormalized = (ActivePercent/100) / overallFractionActive,
         VotedNormalized  = (VotedPercent/100)  / overallFractionVoted)
```

```{r}
normalizedPlotData <- 
  normalized                                                   %>%
  select(Age, 
         `Active Voters` = ActiveNormalized, 
         `Ballots Cast`  = VotedNormalized)                   %>%
  pivot_longer(-Age, names_to = "Group", values_to = "Ratio")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voters", "Ballots Cast"))) %>%
  arrange(Group, Age)
```

```{r NormalizedActiveVoted, fig.width = 8, fig.height = 6}
normalizedVoted <-
  ggplot(normalizedPlotData, aes(Age, y= Ratio, color = Group))  +
  geom_line(size = 2)                                            +
  geom_hline(yintercept = 1.0, color = "blue", 
             linetype = "dotted", size = 1)                      +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)                    +
  scale_y_continuous(breaks = 0.01 * 0:24 * 10)                  +
  scale_color_manual(values = c("black", "red"))                 +
  labs(
        title = paste0(FigureTitle, ":  Normalized Voter Fraction",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Normalized Fraction of Voters",
        caption = c(Source, Stamp)
      )                                                          +
  theme(title = element_text(size = 12))

print(normalizedVoted)
```                   

# By Party

```{r}
ageDataByParty <- ageDataByGroup(voterList, Party)
nrow(ageDataByParty)
```

```{r}
byParty <- bySummary(voterList, Party)
byParty %>%  Show()
```

```{r}
countPlotDataByParty <- 
  ageDataByParty                                                %>%
  select(Age, Party,
         `Registered Voters` = Registered, 
         `Active Voters`     = Active, 
         `Ballots Cast`      = Voted)                           %>%     #####
  pivot_longer(c(-Age,-Party), names_to = "Group", values_to = "Voters")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Registered Voters",
                                   "Active Voters",
                                   "Ballots Cast"))) %>%
  arrange(Group, Age, Party)

nrow(countPlotDataByParty)
```

## Voter Registration Counts

```{r}
partyCounts <- table(voterList$Party)
partyCounts
```


```{r}
partyLabel <- c("D" = paste("Democratic",   comma(partyCounts["D"])),
                "L" = paste("Libertarian",  comma(partyCounts["L"])), 
                "R" = paste("Republican",   comma(partyCounts["R"])), 
                "U" = paste("Unaffiliated", comma(partyCounts["U"])))
```

```{r}
if (length(partyCounts) == 5)   # 2010 and earlier
{
  partyLabel = c(partyLabel[1],
                 "F" = paste("Reform",  comma(partyCounts["F"])),
                 partyLabel[2:4])
}
```

```{r}
registeredActiveVotedByParty <-
  ggplot(countPlotDataByParty, 
         aes(x = Age, y= Voters, color = Group))         +
  geom_line(size = 2)                                    +
  scale_y_continuous(labels = scales::comma)             +     
  scale_x_continuous(breaks = 20 + 0:8 * 10)             +
  scale_color_manual(values = c("grey", "black", "red")) + 
  labs(
        title = paste0(FigureTitle, ":  Voters by Party and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Number of Voters by Age",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank())            +
  facet_wrap(~Party, 
             ncol = length(partyLabel),
             labeller = labeller(Party = partyLabel))    +
  geom_text(data = byParty,
            aes(x = 15,
                y = Inf,
                label = paste0(comma(Active), " Active Voters\n",
                               comma(Voted),  " Ballots Cast")),
            color = "black", hjust = 0, vjust = 1, size = 4)  
```

```{r RegisteredActiveVotedByParty, fig.width = 12, fig.height = 6}     
print(registeredActiveVotedByParty)
```

## Fraction of Voters

```{r}
partyLabelFraction <-
  c("D" = paste0("Democratic ",   
                 round(100 * partyCounts["D"] / sum(partyCounts), 2), "%"),
    
    "L" = paste0("Libertarian ",  
                 round(100 * partyCounts["L"] / sum(partyCounts), 2), "%"),
    
    "R" = paste0("Republican ",   
                 round(100 * partyCounts["R"] / sum(partyCounts), 2), "%"), 
    
    "U" = paste0("Unaffiliated ", 
                 round(100 * partyCounts["U"] / sum(partyCounts), 2), "%"))
```

```{r}
if (length(partyCounts) == 5)   # 2010 and earlier
{
  partyLabelFraction = c(partyLabelFraction[1],
                         "F" = paste0("Reform ",
                                      round(100 * partyCounts["F"] / sum(partyCounts), 2), "%"),
                         partyLabelFraction[2:4])
}
```

```{r}
fractionByPartyAge <-
  ageDataByParty                                                      %>%
  select(Age, Party,
         `Active Voters` = ActivePercent,
         `Ballots Cast`  = VotedPercent)                               %>%
  pivot_longer(c(-Age, -Party), names_to = "Group", values_to = "Percent")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voters", "Ballots Cast")))  %>%
  arrange(Group, Age, Party)
```                 

```{r}
fractionByParty <-
  ageDataByParty
```

```{r}
activeVotedByParty <-
  ggplot(fractionByPartyAge, aes(Age, y= Percent, color = Group)) +
  geom_line(size = 2)                                    +
  scale_y_continuous(breaks =      0:10 * 10)            +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)            +
  scale_color_manual(values = c("black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  Percentage of Registered Voters by Party and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Percentage of Registered Voters",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())            +
  
  facet_grid(col = vars(Party),
             labeller = labeller(Party = partyLabelFraction),
             scales = "free_x")                          +
  
  geom_hline(data = byParty,
             aes(yintercept = ActivePercent), 
             color = "black", 
             linetype = "dotted", size = 1)              +
  geom_text(data = byParty,
            aes(x = 15,
                y = ActivePercent,
                label = paste0(round(ActivePercent, 1), "%")),
            color = "black", hjust = 0, vjust = 1, size = 4)   +
  
  geom_hline(data = byParty,
             aes(yintercept = VotedPercent), 
             color = "red", 
             linetype = "dotted", size = 1)               +
  geom_text(data = byParty,
            aes(x = 15,
                y = VotedPercent,
                label = paste0(round(VotedPercent, 1), "%")),
            color = "red", hjust = 0, vjust = 1, size = 4)  
```

```{r ActiveVotedPercentageByParty, fig.width = 12, fig.height = 6}
print(activeVotedByParty)
```
                               
# By Gender

```{r}
ageDataByGender <- ageDataByGroup(voterList, Gender)
nrow(ageDataByGender)
```

```{r}
byGender <- bySummary(voterList, Gender)
byGender %>% Show()
```

```{r}
countPlotDataByGender <-
  ageDataByGender                                               %>%
  select(Age, Gender,
         `Registered Voters` = Registered,
         `Active Voters`     = Active,
         `Ballots Cast`      = Voted)                           %>%     #####
  pivot_longer(c(-Age,-Gender), names_to = "Group", values_to = "Voters")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Registered Voters",
                                   "Active Voters",
                                   "Ballots Cast"))) %>%
  arrange(Group, Age, Gender)

nrow(countPlotDataByGender)
```

## Voter Registration Counts

```{r}
genderCounts <- table(voterList$Gender)
genderCounts
```


```{r}
genderLabel <- c("Female"  = paste("Female",  comma(genderCounts["Female"])),
                 "Male"    = paste("Male",    comma(genderCounts["Male"])),
                 "Unknown" = paste("Unknown", comma(genderCounts["Unknown"])))
```

```{r}
registeredActiveVotedByGender <-
  ggplot(countPlotDataByGender,
         aes(x = Age, y= Voters, color = Group))         +
  geom_line(size = 2)                                    +
  scale_y_continuous(labels = scales::comma)             +
  scale_x_continuous(breaks = 20 + 0:8 * 10)             +
  scale_color_manual(values = c("grey", "black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  By Gender and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Number of Voters by Age",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank())            +
  facet_wrap(~Gender,
             ncol = length(genderLabel),
             labeller = labeller(Gender = genderLabel))    +
  geom_text(data = byGender,
            aes(x = 15,
                y = Inf,
                label = paste0(comma(Active), " Active Voters\n",
                               comma(Voted),  " Ballots Cast")),
            color = "black", hjust = 0, vjust = 1, size = 4)
```

```{r RegisteredActiveVotedByGender, fig.width = 12, fig.height = 6}
print(registeredActiveVotedByGender)
```

## Fraction of Voters

```{r}
genderLabelFraction <-
  c("Female" = paste0("Female ",
                 round(100 * genderCounts["Female"]  / sum(genderCounts), 2), "%"),

    "Male" = paste0("Male ",
                 round(100 * genderCounts["Male"]    / sum(genderCounts), 2), "%"),

    "Unknown" = paste0("Unknown ",
                 round(100 * genderCounts["Unknown"] / sum(genderCounts), 2), "%"))
```

```{r}
fractionByGenderAge <-
  ageDataByGender                                                     %>%
  select(Age, Gender,
         `Active Voter` = ActivePercent,
         `Ballots Cast` = VotedPercent)                               %>%
  pivot_longer(c(-Age, -Gender), names_to = "Group", values_to = "Percent")  %>%
  mutate(Group = factor(Group,  # control order for legend
                        levels = c("Active Voter", "Ballots Cast")))  %>%
  arrange(Group, Age, Gender)
```

```{r}
fractionByGender <-
  ageDataByGender
```

```{r}
activeVotedByGender <-
  ggplot(fractionByGenderAge, aes(Age, y= Percent, color = Group)) +
  geom_line(size = 2)                                    +
  scale_y_continuous(breaks =      0:10 * 10)            +
  scale_x_continuous(breaks = 20 + 0: 8 * 10)            +
  scale_color_manual(values = c("black", "red")) +
  labs(
        title = paste0(FigureTitle, ":  Percentage By Gender and Age",
                       " - ", FigureDistrict),
        x = "Election Day Age[years]",
        y = "Percentage of Registered Voters",
        caption = c(Source, Stamp)
      )                                                  +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())            +

  facet_grid(col = vars(Gender),
             labeller = labeller(Gender = genderLabelFraction),
             scales = "free_x")                          +

  geom_hline(data = byGender,
             aes(yintercept = ActivePercent),
             color = "black",
             linetype = "dotted", size = 1)              +
  geom_text(data = byGender,
            aes(x = 15,
                y = ActivePercent,
                label = paste0(round(ActivePercent, 1), "%")),
            color = "black", hjust = 0, vjust = 1, size = 4)   +

  geom_hline(data = byGender,
             aes(yintercept = VotedPercent),
             color = "red",
             linetype = "dotted", size = 1)               +
  geom_text(data = byGender,
            aes(x = 15,
                y = VotedPercent,
                label = paste0(round(VotedPercent, 1), "%")),
            color = "red", hjust = 0, vjust = 1, size = 4)
```

```{r ActiveVotedPercentageByGender, fig.width = 12, fig.height = 6}
print(activeVotedByGender)
```

# By Precinct

```{r}
ageDataByPrecinct <- ageDataByGroup(voterList, Precinct)
nrow(ageDataByPrecinct)
```

```{r}
byPrecinct <- 
  bySummary(voterList, Precinct)  %>%
  rename(OverallActivePercent = ActivePercent,
         OverallVotedPercent  = VotedPercent)
byPrecinct %>%  Show(bigMark = ",") %>%  scroll_box(height = "300px")
```

## Precinct Size

```{r Precinct_Density}
ggplot(byPrecinct, aes(x = Registered, y=..density..))          + 
  geom_histogram(fill = "skyblue", color = "grey80", bins = 50) +
  scale_x_continuous(labels = scales::comma)                    +
  labs(title = paste("Precinct Sizes in", FigureDistrict),
       subtitle = paste("All", nrow(byPrecinct), "precincts"),
       x = "Precinct Size [registered voters]",
       y = "Count",
       caption=c(Source, Stamp))                                   
```

```{r}
largeEnough   <- 
  byPrecinct  %>%
  filter(Registered >= PRECINCT_MINIMUM)  

nrow(largeEnough)
```
  
```{r Precinct_Density_LargeEnough}
ggplot(largeEnough, aes(x = Registered, y=..density..))          + 
  geom_histogram(fill = "skyblue", color = "grey80", bins = 50)  +
  scale_x_continuous(labels = scales::comma)                     +
  labs(title = paste("Precinct Sizes in", FigureDistrict),
       subtitle = paste0(nrow(largeEnough), " precincts",
                      " [ignoring ", nrow(byPrecinct) - nrow(largeEnough), 
                      " precincts with fewer than ", PRECINCT_MINIMUM, " registered voters]"),
       x = "Precinct Size [registered voters]",
       y = "Count",
       caption=c(Source, Stamp))                                   
  
```

For now, let's only look at voter turnout.

```{r}
normalizedAgeDataByPrecinct <-
  byPrecinct                                   %>%
  select(Precinct, OverallVotedPercent)        %>%
  inner_join(ageDataByPrecinct,
             by = "Precinct")                  %>%
  mutate(NormalizedFractionVoted = VotedPercent / OverallVotedPercent)  %>%
  select(Precinct, Age, NormalizedFractionVoted) %>%
  arrange(Precinct, Age)                         

dim(normalizedAgeDataByPrecinct)
```

## Voter Registration Counts

```{r}
precinctCounts <- table(voterList$Precinct)
```

```{r}
sum(precinctCounts)
```

## Scatterplot of County Overall with All Precincts

```{r}
countyNormalized <- 
  normalizedPlotData                       %>%
  rename(NormalizedFractionVoted = Ratio)  %>%
  filter(Group == "Ballots Cast")          %>%
  select(-Group)                           %>%
  mutate(Precinct = params$DistrictName)   %>%
  relocate(Precinct, .before = Age)         

dim(countyNormalized)
```

```{r NormalizedVotedTurnout, fig.width = 8, fig.height = 6}
normalizedVoted <-
  ggplot()                                      +
  
  geom_line(data = semi_join(normalizedAgeDataByPrecinct, 
                             largeEnough,
                             by = "Precinct"),
            aes(x = Age, y= NormalizedFractionVoted, 
                group = Precinct),
            color = "grey", size = 1)           +
  
  geom_line(data = countyNormalized, 
            aes(x = Age, y= NormalizedFractionVoted),
            color = "red", size = 2)            +
  
  geom_hline(yintercept = 1.0, color = "blue", 
             linetype = "dotted", size = 1)     +
  
  geom_text(aes(x = 85,
                y = 0.1,
                label = "Low counts cause 'noise' for many ages"),
            color = "black", hjust = 1, vjust = 1, size = 4)   +
  
  scale_x_continuous(breaks = 2:10 * 10)        +
  scale_y_continuous(breaks = 0:10 * 0.2,
                     limits = c(0, 2.0) )       +
  
  labs(
        title = paste0(FigureTitle, ":  Normalized Voter Faction by Age"),
        subtitle = paste0(FigureDistrict, " = red line; ",
                          nrow(largeEnough), " Precincts = grey lines;"),
        x = "Election Day Age[years]",
        y = "Normalized Fraction of Voters",
        caption = c(Source, Stamp)
      )                                         +
  theme(
         panel.grid.minor.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         title = element_text(size = 12)
       )                        

print(normalizedVoted)
```

## Heatmap

Create matrix

```{r}
County <- 
  bind_rows(countyNormalized,
            semi_join(normalizedAgeDataByPrecinct, 
                             largeEnough,
                             by = "Precinct"),)  %>%
  pivot_wider(names_from = Age,
              values_from = NormalizedFractionVoted,
              values_fill = list(NormalizedFractionVoted = 0))  %>%
  mutate(across(everything(), ~replace_na(.x, 0)))   # kludge fix (e.g., Primary 2006)    
  
```

```{r}
comparisonMatrix <- as.matrix(County %>% select(-Precinct))
dim(comparisonMatrix)
```

```{r}
dimnames(comparisonMatrix) <- list(
                                    County$Precinct,
                                    colnames(County[-1]) %>% as.numeric()
                                   )
```

```{r}
ratioStats <- fivenum(comparisonMatrix)

ratioStats
```

```{r}
colorBreaks <-
  c(
    seq(0.0, 1.0, length.out = 12),
    seq(1.0, 2.0, length.out = 6)[-1]
  )

colorBreaks
```

Attempt to "balance" palette above and below 1.

```{r}
colorPalette <-
  c(colorRampPalette(c("red",   "white"))(12)[1:11],    # skip last
    "#FFFFFF",                                        # color for ratio of 1
    colorRampPalette(c("white", "skyblue"))(5)[2:5])  # skip first
    
colorPalette
```

```{r}
# layout matrix 1 = heatmap, 2 = row dendrogrm, 3 = column dendrogram, 4 = key
lmat <- rbind( c(0, 3, 3),  # layout matrix
               c(2, 1, 1),
               c(0, 0, 4))

lwid <- c(3, 4, 4)          # entry for each column in layout matrix

# relative height of the rows. must have length = nrow(lmat)
lhei <- c(0.25, 4, 0.25)    # entry for each row in layout matrix
```

```{r heatmap, fig.width = 20, fig.height = 32}
oldPar <- par(mar = c(5, 0, 2, 16))      

heatmap.2(
           comparisonMatrix,
           scale      = "none",
           dendrogram = "row",
           Colv       = NA,         # suppress reordering of columns
           
           trace  = "none",         # turns off trace lines inside the heat map
           col    = colorPalette,
           breaks = colorBreaks,
           
           margins = c(4, 16),      # column and row names
           
           cexRow = 1,    
           cexCol = 1.25,
           
           key.title = NA,           # suppress title
           key.xlab  = "Normalized Turnout",
           keysize   = 3,
           density.info = "none",    # turns off density plot inside color legend
          
           lmat = lmat, lwid = lwid, lhei = lhei
         )

title(main = paste0(FigureTitle, 
                    ":  Normalized Voter Faction by Age - ",
                    FigureDistrict),
      cex.main = 2,
      sub = paste(Source, " - ", Stamp), 
      cex.sub  = 1.5)

par(oldPar)  
```

# Fini {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r}
sessionInfo()
```

</div>

```{r ThatsAll, echo = FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`
                                                                                         