---
title:  "Driver to Create County Turnout Reports"
author: "Earl F Glynn"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
editor_options: 
  chunk_output_type: console
---

```{r Dsetup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r Dsetup2, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment    = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

##  {.active}

## Constants

```{r DConstants}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r DPackages}
library(rmarkdown)
library(tidyverse)
library(lubridate)

library(readxl)
library(kableExtra) 

library(foreach)
library(parallel)
library(doParallel)       
```


## Helper Functions

```{r DHelper}
Show <- function(data, caption="", bigMark="", ...)
{
  data                                                    %>%
  kable("html", 
        caption = caption,
        format.args = list(big.mark = bigMark))   %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"),
                position = "left", 
                full_width = FALSE, ...)
}
```

Read `efg`s "standardized" preprocessed version of the voter file from the Kansas Secretary of State.

```{r DReadVoterFileFunction}
readVoterFile <- function(Filename, ELECTION_DATE)
{
  voterFile <-
  read_csv(Filename, guess_max = EXCEL_LIMIT) %>%
  select(
          ID14,                   # Hash of County, Name, DOB, Gender
          SOSKey,
          Status,
          BirthDate,
          RegisterDate,
          Gender,
          County, City,
          Precinct1 = Precinct,
          Precinct2,
          Party,                 # D, R, L, U, ?
          LastVoteYear,
          TotalVotes,            # Number of ballots in  voting history (max is 10)
          August,                # Number of August ballots cast in even years
          AugPattern,            # Right most "X" is for August 2020 Primary
          November,              # Number of November ballots cast in even years
          NovPattern             # Right most "X" is for November 2020 General
        )             %>%
  mutate(Gender       = recode(Gender,
                               "F"      = "Female",
                               "M"      = "Male",
                               "?"      = "Unknown",
                               .missing = "Unknown"),
         Party        = recode(Party,
                               "?"      = "U",
                               .missing = "U"),
         Precinct     = paste0(Precinct1, " - ", Precinct2),
         BirthDate    = as.Date(BirthDate,    "%m/%d/%Y"),
         RegisterDate = as.Date(RegisterDate, "%m/%d/%Y"),
         Age          = round((BirthDate %--% as.Date(ELECTION_DATE)) /
                              years(1)),
         Registered   = (RegisterDate %--% as.Date(ELECTION_DATE)) /
                         years(1),
         RegisteredTooLate = as.integer(
                                         as.Date(RegisterDate, "%m/%d/%Y") >
                                         (as.Date(ELECTION_DATE) - 21)
                                       )                       
        )             %>%
    select(-Precinct1, - Precinct2)

  invisible(voterFile)
}
```

# Metadata

```{r DShow}
metadata <- read_xlsx("../Statewide/Statewide-Elections.xlsx")
nrow(metadata)
```

Process only one

```{r DSelectOne}
info <- metadata %>% slice(7)
info  %>% Show()
```                              

# Parameters

Use first publicly available voter list after a general election whenever possible.

```{r DParms}
PRECINCT_MINIMUM_VOTERS <-  25
```

# Read Kansas Data

```{r DReadVoterFile}
afterFilename <- paste0("E:/Data/Kansas/RawData/",
                        str_sub(info$ELECTION_DATE, 1, 3), "0s/",     #####
                        str_sub(info$AFTER_FILE_DATE, 1, 4), "/",
                        info$AFTER_FILE_DATE, "/KSCVRID.CSV")
VOTER_MASTER <- readVoterFile(afterFilename, info$ELECTION_DATE)  
```


```{r DAddVited}
if (info$ELECTION_NAME == "Primary")
{
  VOTER_MASTER <-
    VOTER_MASTER %>%
    mutate(
            VotedInElection =
              as.integer(substr(AugPattern, info$HISTORY_INDEX, info$HISTORY_INDEX) == "X")
          )
}

if (info$ELECTION_NAME == "General")
{
  VOTER_MASTER <-
    VOTER_MASTER %>%
    mutate(
            VotedInElection =
              as.integer(substr(NovPattern, info$HISTORY_INDEX, info$HISTORY_INDEX) == "X")
          )
}
```                                 

```{r DSummary}
VOTER_MASTER %>%
  summarize(
             n         = n(),
             nID14     = n_distinct(ID14),
             nSOSKey   = n_distinct(SOSKey),
             nVoted    = sum(VotedInElection, na.rm = TRUE),
             nParty    = n_distinct(Party),
             nGender   = n_distinct(Gender),
             nCounty   = n_distinct(County),
             nPrecinct = n_distinct(Precinct),
             minAge    = min(Age, na.rm = TRUE),
             maxAge    = max(Age, na.rm = TRUE)
            )  %>%
  Show(bigMark = ",")
```

```{r DCounties}
counties <- sort(unique(VOTER_MASTER$County))
length(counties)
```

# Function to Process Single County

```{r DProcessCounty}
processCounty <- function(county, info, VOTER_MASTER)
{
  title <- paste0("Normalized Turnout for ",
                  county, " County, KS in ",
                  info$ELECTION_NAME, " Election ",
                  info$ELECTION_DATE)     
  
  OUTPUT_DIR <- paste0("./", info$ELECTION_DATE, "/", county)      

  rmarkdown::render(
                     "County-Normalized-Turnout-Template.Rmd",
                     output_format = html_document(theme      = "cerulean",
                                                   toc        = TRUE,
                                                   toc_format = TRUE,
                                                   toc_float  = TRUE,
                                                   number_sections = TRUE,
                                                   code_folding = "hide"),   

                     output_file = paste0(info$ELECTION_DATE, "-", county, "-County.html"),
                     output_dir  = OUTPUT_DIR,

                     quiet  = TRUE,
                     params = list(
                                    ReportTitle     = title,
                                    District        = "County",
                                    DistrictName    = county,
                                    ElectionDate    = info$ELECTION_DATE,
                                    ElectionName    = info$ELECTION_NAME,
                                    AfterFileDate   = info$AFTER_FILE_DATE,
                                    PrecinctMinimum = PRECINCT_MINIMUM_VOTERS,
                                    voterMaster     = VOTER_MASTER
                                  )
                   )

  # Add election date and county name to all graphics files
  listFiles <- dir(path = ".",  pattern = "^.*\\.png$|Too.*\\.xlsx$")    
  newNames <- paste0(info$ELECTION_DATE, "-", county, "-", listFiles)
  file.rename(listFiles, newNames)

  #Brute force to get all files into `output_dir`
  listFiles <- dir(path = ".", pattern = "^.*\\.png$|Too.*\\.xlsx$")           
  file.copy(listFiles, OUTPUT_DIR, copy.date = TRUE)
  file.remove(listFiles)   
  
  county
}
```

# Process All Counties

```{r DProcessAll}
system.time(
             for (county in counties)
             { 
               processCounty(county, info, VOTER_MASTER)
             }
           )
```
                          

# Fini {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r Dsession}
sessionInfo()
```

</div>

```{r ThatsAllDriver, echo = FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

